Motional Quantum Ground-State Cooling Outside the Lamb-Dicke Regime

### 2017-06-28 paper.tex
Supplemental material about simulation
Rabi flopping reference, heating/loss number

Taking all these features into account, we use a Monte-Carlo simulation to verify
the validity of our method \fxnote{WRITE BRIEF SUMMARY IN APPENDIX?}.
In the simulation, we can observe the high heating rate due to the high LD parameters
and confirm that by using high order Raman sideband transition in the cooling sequence we can
suppress this effect and reduces the motional energy of the atom faster.
The simulation is also used to guide the optimization of the cooling sequence by exploring the
large parameter space and finding a robust cooling strategy.
As shown in figure \ref{f-setup}C, we found that instead of cooling on only one sideband order
at a time, it is generally more efficient to alternate the cooling pulse between two
neighboring orders (axial) to minimize the accumulation of atom in a state
not addressed by a particular order.
The simulation also shows that an efficient way of cooling on a high order sideband is
to set the Rabi frequency of the Raman transition to perform a $\pi$ pulse for the maximum matrix element of that order over all states (i.e. the maxima in \ref{f-ld}A).
This makes sure we are using each order to cool a range of states that has the strongest
coupling to this order and let other sidebands to cool other states.

@@

It is important to note that the large Raman LD parameters $\eta^R$ mean that
we must use caution when measuring the temperature.
Traditional sideband thermometry uses
the ratio of the cooling and heating sidebands to measure $\bar n / (\bar n + 1)$, which relies
on the coupling strength to be proportional to $\sqrt{n}$. However, outside the
LD regime, the coupling strength deviates from this simple scaling rule rapidly as
shown in figure \ref{f-ld}A.
When the atom temperature is still high,
we measure the heights of multiple sidebands to make sure no population is hidden at the
minimum of one sideband order. Once the atom is cooled down, we estimate the temperature
and ground state population under the assumption that only a few states are populated.
We can verify the result via an independent measurement of Rabi flopping on the carrier and heating
sidebands. This data is shown before and after cooling in Figures
\ref{f-radial}B and \ref{f-radial}C and yields $89(2)\%$ and $92(2)\%$
showing good agreement between the two methods.

The axial direction is much less confined and therefore has a much higher initial effective
LD parameter of $\eta^{OP}_{eff1}\approx 4.0$, and is therefore more difficult to cool.
The red line in figure \ref{f-axial}A shows the Raman spectrum in the axial direction
before applying Raman sideband cooling; the clearly resolved sidebands
up to eighth order indicates population in highly-excited motional levels.
Therefore, we begin our cooling sequence by driving Raman transitions on the eighth order axial
cooling sideband. After applying the cooling sequence identical to the one we use to cool
the radial directions, the spectrum is shown in figure \ref{f-axial}A as the blue line.
All of the high order ($\geqslant2$) axial cooling sidebands disappear, and the first order
cooling sideband is strongly suppressed.
The ground state probability calculated from this spectrum is $92(3)\%$.
We can again use Rabi flopping on the carrier and heating sideband to verify this result
similar to the radial direction. In this case, we see very good agreement on the carrier
(figure \ref{f-axial}B) but there is additional decoherence on the axial first order
heating sideband (figure \ref{f-axial}C).
We believe this decoherence is caused by technical noise, which is the most pronounced
on this transition due to its slow Rabi frequency.
The decoherence time scale is consistent with magnetic field fluctuations of $1.5$ mG that we have measured in the lab, which produces
a Zeeman shift of around $3$ kHz.

Combining the axial and radial cooling results,
we have achieved a 3D ground state probability of $77(4)\%$.
We believe this is currently limited by off-resonant scattering from the Raman beams
and the resonance shift caused by magnetic field fluctuations.
We are planing to improve these by increasing the detuning of the Raman beams
and implementing better control of the magnetic field.
We expect these changes to improve the cooling performance even further.
We have shown that despite the difficulty in cooling sodium atoms,
it is possible to achieve reliable three dimensional cooling with significant ground state
probability using new cooling techniques and a optimized cooling sequence.
This opens up the possibility of cooling a large variety of systems to their motional ground
state using Raman sideband cooling that is challenging to laser cool
including molecules and exotic atoms.

>>>>

Taking all these features into account, we use a Monte-Carlo simulation to verify
the validity of our method.
In the simulation, we can observe the high heating rate due to the high LD parameters
and confirm that by using high order Raman sideband transition in the cooling sequence we can
suppress this effect and reduces the motional energy of the atom faster.
The simulation is also used to guide the optimization of the cooling sequence by exploring the
large parameter space and finding a robust cooling strategy.
As shown in figure \ref{f-setup}C, we found that instead of cooling on only one sideband order
at a time, it is generally more efficient to alternate the cooling pulse between two
neighboring orders (axial) to minimize the accumulation of atom in a state
not addressed by a particular order.
The simulation also shows that an efficient way of cooling on a high order sideband is
to set the Rabi frequency of the Raman transition to perform a $\pi$ pulse for the maximum matrix element of that order over all states (i.e. the maxima in \ref{f-ld}A).
This makes sure we are using each order to cool a range of states that has the strongest
coupling to this order and let other sidebands to cool other states.

@@

It is important to note that the large Raman LD parameters $\eta^R$ mean that
we must use caution when measuring the temperature.
Traditional sideband thermometry uses
the ratio of the cooling and heating sidebands to measure $\bar n / (\bar n + 1)$, which relies
on the coupling strength to be proportional to $\sqrt{n}$. However, outside the
LD regime, the coupling strength deviates from this simple scaling rule rapidly as
shown in figure \ref{f-ld}A.
When the atom temperature is still high,
we measure the heights of multiple sidebands to make sure no population is hidden at the
minimum of one sideband order. Once the atom is cooled down, we estimate the temperature
and ground state population under the assumption that only a few states are populated.
We can verify the result via an independent measurement of Rabi flopping on the carrier and
heating sidebands\cite{Meekhof1996}. This data is shown before and after cooling in Figures
\ref{f-radial}B and \ref{f-radial}C.
The data is fitted using same Monte-Carlo simulation we used to simulate the cooling
which yields $89(2)\%$ and $92(2)\%$ showing good agreement between the two methods.

The axial direction is much less confined and therefore has a much higher initial
$\bar n_1=21(2)$ and effective LD parameter of $\eta^{OP}_{eff1}\approx 4.0(2)$,
and is therefore more difficult to cool.
The red line in figure \ref{f-axial}A shows the Raman spectrum in the axial direction
before applying Raman sideband cooling; the clearly resolved sidebands
up to eighth order indicates population in highly-excited motional levels.
Therefore, we begin our cooling sequence by driving Raman transitions on the eighth order axial
cooling sideband. After applying the cooling sequence identical to the one we use to cool
the radial directions, the spectrum is shown in figure \ref{f-axial}A as the blue line.
All of the high order ($\geqslant2$) axial cooling sidebands disappear, and the first order
cooling sideband is strongly suppressed.
The ground state probability calculated from this spectrum is $92(3)\%$.
We can again use Rabi flopping on the carrier and heating sideband to verify this result
similar to the radial direction. In this case, we see very good agreement on the carrier
(figure \ref{f-axial}B) but there is additional decoherence on the axial first order
heating sideband (figure \ref{f-axial}C).
We believe this decoherence is caused by technical noise, which is the most pronounced
on this transition due to its slow Rabi frequency.
The decoherence time scale is consistent with magnetic field fluctuations of $1.5$ mG that we have measured in the lab, which produces
a Zeeman shift of around $3$ kHz.

Combining the axial and radial cooling results,
we have achieved a 3D ground state probability of $77(4)\%$.
We believe this is currently limited by off-resonant scattering from the Raman beams,
which are measured to be between $3$ to $15$ kHz for the detuning and power we use,
and the resonance shift caused by magnetic field fluctuations.
Other imperfections of the current experiement include a $5\%$ loss during the single atom
imaging and an additional $10\%$ due to the heating from the Raman beams during cooling.
We are planing to improve these by increasing the detuning of the Raman beams,
implementing better control of the magnetic field and optimize the temperature during imaging.
We expect these changes to improve the cooling performance even further.
We have shown that despite the difficulty in cooling sodium atoms,
it is possible to achieve reliable three dimensional cooling with significant ground state
probability using new cooling techniques and a optimized cooling sequence.
This opens up the possibility of cooling a large variety of systems to their motional ground
state using Raman sideband cooling that is challenging to laser cool
including molecules and exotic atoms.
